### 2023-04-11 analyses.Rmd
- initalize repo

---
title: "Analyses"
bibliography: 
  - [bibliography/references.bib]
output: 
  html_document:
    toc: TRUE
    toc_float:
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
---

```{r setup, include=F}
knitr::opts_chunk$set(
  cache = TRUE, 
  warning = FALSE, 
  echo = TRUE, 
  error = FALSE, 
  message = FALSE, 
  prompt = FALSE, 
  tidy = TRUE, 
  fig.height = 6, 
  fig.width = 8
  )

options(
  digits = 3, 
  width = 120
  )
```

In what follows, you can find the analyses for the results reported in the paper. To see the underlying code, click on the button "code".

# Setup
## Packages

```{r results='hide', cache=F}
# install github packages
# devtools::install_github("tdienlin/td@v.0.0.2.3")  # uncomment to install
# devtools::install_github("yrosseel/lavaan")

# define packages
packages <- 
  c(
    "brms", 
    "cowplot", 
    "devtools", 
    "english", 
    "faoutlier", 
    "GGally", 
    "gganimate",
    "kableExtra", 
    "knitr", 
    "lavaan", 
    "magrittr",
    "marginaleffects",
    "MVN", 
    # "psych", 
    "pwr", 
    "quanteda", 
    "semTools", 
    "tidyverse", 
    "td"
    )

# load packages
lapply(
  packages, 
  library, 
  character.only = TRUE, 
  quietly = TRUE
  )

# If not wanting to compute all transformations etc., simply load workspace
# load("data/workspace.RData")
```

## Custom Functions

```{r}
# define function to silence result output
# needed for brms
hush <- 
  function(
    code
    ){
    sink("/dev/null") # use /dev/null in UNIX
    tmp = code
    sink()
    return(tmp)
    }

# plot distributions of items
plot_dist <-
  function(
    item,
    n_rows = 1
    ) {
    ggplot(
      gather(
        select(
          d, 
          starts_with(
            item
            )
          )
        ), 
      mapping = aes(
        x = value
        )
      ) +
      geom_bar() +
      facet_wrap(
        ~key, 
        nrow = n_rows
        ) + 
      theme_bw()
  }

get_desc <- 
  function(
    item,
    name
  ) {
    desc <- 
      select(
        d, 
        starts_with(
          item
          )
        ) %>% 
      apply(
        1, 
        mean, 
        na.rm = TRUE
        ) %>% 
      as.data.frame() %>% 
      summarise(
        m = mean(., na.rm = TRUE), 
        sd = sd(., na.rm = TRUE)
        )
    assign(
      paste0(
        "des_",
        name
      ),
      desc,
      envir = .GlobalEnv
    )
  }

# Function to display diagnostics of fitted hurdle models
model_diagnostics <- 
  function(
    model
  ){
  plot_grid(
    pp_check(
      model,
      type = "dens_overlay",
      nsamples = 100
    ),
    pp_check(
      model,
      type = "loo_pit_qq",
      nsamples = 100
    ),
    pp_check(
      model,
      type = "stat",
      stat = "median",
      nsamples = 100
    ),
    pp_check(
      model,
      type = "stat",
      stat = "mean",
      nsamples = 100
    ),
    labels = c("Density overlay", "LOO-PIT QQ", "Predicted medians", "Predicted means"),
    ncol = 2,
    label_size = 8,
    hjust = 0,
    vjust = 0,
    label_x = 0,
    label_y = 0.93
  )
  }

# wrapper function that runs relevant analyses for hurdle models
run_hurdles <- 
  function(
    object,
    name,
    outcome,
    predictor,
    x_lab,
    y_lab,
    plot = FALSE,
    ...
    ){

    # define define general model
    formula_mdl <- 
      formula(
        paste0(
          outcome,
          " ~ 1 + age + male + edu +",
          predictor
        )
    )

    # define hurdle model
    formula_hrdl <- 
      formula(
        paste0(
          "hu ~ 1 + age + male + edu +",
          predictor
        )
    )

    # fit model
    fit <- 
      hush(
        brm(
          bf(
            formula_mdl, 
            formula_hrdl
            ), 
          data = object, 
          family = hurdle_gamma(),
          silent = TRUE,
          refresh = 0
          )
      )

    # export fit
    assign(
      paste0(
        "fit_",
        name),
      fit,
      envir = .GlobalEnv
    )

    # plot model
    if(
      isTRUE(plot
             )) {
      plot(fit, ask = FALSE)
      }

    # run diagnostics
    fit_diagnostics <- 
      model_diagnostics(fit)
    plot(fit_diagnostics)

    # print summary
    fit_summary <- 
      summary(fit)
    print(fit_summary)

    # calculate slopes
    slopes <-
      cbind(
        Outcome = outcome,
        avg_slopes(fit)
      )

    # print slopes
    cat("\nResults of marginal effects:\n")
    print(slopes)

    # export as object to environment
    assign(
      paste0(
        "slopes_",
        name
        ),
      slopes,
      envir = .GlobalEnv
    )

    # make figure
    fig_res <- 
      conditional_effects(
        fit
        , ask = FALSE
      )

    # make figure
    fig <- 
      plot(
        fig_res,
        plot = FALSE
      )[[predictor]] +
      labs(
        x = x_lab,
        y = y_lab
        )

    # plot figure
    print(fig)

    # export as object to environment
    assign(
      paste0(
        "fig_",
        name
      ),
      fig,
      envir = .GlobalEnv
    )
}
```

## Data-Wrangling

```{r results='hide', cache=F}
# load data
# please note that in a prior step several different data sets were merged using information such as case_taken or IP-addresses. To guarantee the privacy of the participants, these data were deleted after merging.
d_raw <- read_csv("data/data_raw.csv")

# recode variables
d <- 
  d_raw %>% 
  dplyr::rename(
    "male" = "SD01", 
    "age" = "SD02_01", 
    "state" = "SD03", 
    "edu_fac" = "SD04",
    "TR02_01" = "TR01_01",
    "TR02_02" = "TR01_05",
    "TR02_03" = "TR01_09",
    "TR01_01" = "TR01_02",
    "TR01_02" = "TR01_03",
    "TR01_03" = "TR01_04",
    "TR01_04" = "TR01_06",
    "TR01_05" = "TR01_07",
    "TR01_06" = "TR01_08",
    "TR01_07" = "TR01_10",
    "TR01_08" = "TR01_11",
    "TR01_09" = "TR01_12"
    ) %>% 
  mutate(

    #general
    version = factor(
      VE01_01, 
      labels = c(
        "Control", 
        "Like", 
        "Like & Dislike"
        )
      ),
    version_lkdslk = factor(
      version, 
      levels = c(
        "Like & Dislike", 
        "Like", 
        "Control"
        )
      ),
    version_no = recode(
      version, 
      "Control" = 1, 
      "Like" = 2, 
      "Like & Dislike" = 3
      ),
    male = replace_na(
      male, 
      "[3]") %>% 
      dplyr::recode(
        "männlich [1]" = 1, 
        "weiblich [2]" = 0
        ),
    age_grp = cut(
      age, 
      breaks = c(0, 29, 39, 49, 59, 100), 
      labels = c("<30", "30-39", "40-49", "50-59", ">59")
      ),

    # contrasts
    like = recode(
      version, 
      "Control" = 0, 
      "Like" = 1, 
      "Like & Dislike" = 0
      ),
    likedislike = recode(
      version, 
      "Control" = 0, 
      "Like" = 0, 
      "Like & Dislike" = 1
      ),
    control = recode(
      version, 
      "Control" = 1, 
      "Like" = 0, 
      "Like & Dislike" = 0
      ),

    # recode
    edu = recode(
      edu_fac, 
      "Hauptschulabschluss/Volksschulabschluss" = 1, 
      "Noch Schüler" = 1,
      "Realschulabschluss (Mittlere Reife)" = 1,
      "Abschluss Polytechnische Oberschule 10. Klasse (vor 1965: 8. Klasse)" = 1,
      "Anderer Schulabschluss:" = 1,
      "Fachhochschulreife (Abschluss einer Fachoberschule)" = 2,
      "Abitur, allgemeine oder fachgebundene Hochschulreife (Gymnasium bzw. EOS)" = 2,
                 "Hochschulabschluss" = 3),
    PC01_03 = 8 - PC01_03, 
    SE01_05 = 8 - SE01_05, 
    SE01_06 = 8 - SE01_06,

    # behavioral data
    words = replace_na(
      words, 
      0
      ),
    self_dis = words + (reactions * 2),
    time_read_MIN = time_read / 60,
    time_sum_min_t2 = TIME_SUM_t2 / 60,

    # take logarithm
    words_log = log1p(words)
  )

# variable labels
var_names <- c(
  "Privacy concerns", 
  "General gratifications", 
  "Specific gratifications",
  "Privacy deliberation", 
  "Self-efficacy", 
  "General trust", 
  "Specific trust",
  "Words log"
  )

var_names_breaks <- c(
  "Privacy\nconcerns", 
  "General\ngratifications", 
  "Specific\ngratifications",
  "Privacy\ndeliberation", 
  "Self-\nefficacy", 
  "General\ntrust", 
  "Specific\ntrust",
  "Words (log)"
  )

# Extract sample characteristics and descriptives.
# sample descriptives t1
n_t1 <- nrow(d)
age_t1_m <- mean(d$age, na.rm = TRUE)
male_t1_m <- mean(d$male, na.rm = TRUE)
college_t1_m <- table(d$edu)[3] / n_t1

# sample descriptives t2
# participants
n_t2 <- 
  filter(
    d, 
    !is.na(
      id_survey_t2
      )
    ) %>% 
  nrow()

# age
age_t2_m <- 
  filter(
    d, 
    !is.na(
      id_survey_t2
      )
    ) %>% 
  select(
    age
    ) %>% 
  mean()

# gender
male_t2_m <- 
  filter(
    d, 
    !is.na(
      id_survey_t2
      )
    ) %>% 
  select(
    male
    ) %>% 
  mean(
    na.rm = TRUE
    )

# education
college_t2_m <- 
  filter(
    d, 
    !is.na(
      id_survey_t2
      )
    ) %>% 
  select(
    edu
    ) %>% 
  table() %>% 
  .[3] / n_t2

# descriptives users
n_users <- 
  filter(
    d, 
    !is.na(
      post_count
      )
    ) %>% 
  nrow()

# characteristics of posts
n_comments <- 
  sum(
    d$post_count, 
    na.rm = TRUE
    )

n_words <- 
  sum(
    d$words, 
    na.rm = TRUE
    )

n_time <- 
  sum(
    d$time_read_MIN, 
    na.rm = TRUE
    )

n_posts <- 
  sum(
    d$post_count, 
    na.rm = TRUE
    )

# filter unmatched cases, use only completes
d <- 
  filter(
    d, 
    !is.na(
      post_count
      ) & !is.na(
        id_survey_t2
        )
    )

n_matched <- nrow(d)

# save data file with all participants to compare results
d_all <- d
```

# Filter Participants

We filtered participants who answered the questionnaire in less than three minutes, which we considered to be unreasonably fast.

```{r cache=F}
# filter speeders
time_crit <- 3 # minimum time on survey

# count number of speeders
n_speeding <- 
  nrow(
    filter(
      d, 
      time_sum_min_t2 < time_crit
      )
    )

# Deletion of fast respondents
d <- 
  filter(
    d, 
    time_sum_min_t2 >= time_crit
    )
```

We inspected the data manually for cases with obvious response patterns. The following cases show extreme response patterns (alongside fast response times), and were hence removed.

```{r cache=F}
# identify response patterns
resp_pattern <- c(
  "ANIEVLK9F2SW", 
  "BN4MAOWZO7W2"  # clear response pattern
  )

n_resp_pattern <- 
  length(
    resp_pattern
    )

# infl_cases_tokens <- infl_cases_tokens[!infl_cases_tokens %in% resp_pattern]

d %>% 
  filter(
    case_token %in% resp_pattern
    ) %>% 
  select(
    case_token, 
    GR01_01:SE01_06, 
    topics_entered:reactions, 
    -SO01_01, 
    TIME_SUM_t1, 
    TIME_SUM_t2
    ) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%")

d %<>% 
  filter(
    !case_token %in% resp_pattern
    )

# sample descriptives final data set
n_final <- nrow(d)
age_final_m <- mean(d$age)
male_final_m <- mean(d$male, na.rm = TRUE)
college_final_m <- table(d$edu)[[3]] / n_final
```

# Measures
## Privacy concerns
### Items

Using the participation platform I had ...

 1. ... concerns about what happens to my data.
 2. ... concerns about disclosing information about myself.
 3. ... no concerns. (reversed)
 4. ... concerns that others could discover my real identity (i.e. my last and first name).
 5. ... concerns that information about myself could fall into wrong hands.
 6. ... concerns that others could discover what my political views are.
 7. ... concerns about my privacy.

### Distributions

```{r fig.height = 3}
# plot distribution
plot_dist(
  item = "PC01"
)

# extract results
get_desc(
  item = "PC01",
  name = "pricon"
)
```

### CFA

```{r}
name <- "pricon"
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_03 + PC01_04 + PC01_05 + PC01_06 + PC01_07
"
fit <- 
  lavaan::sem(
    model = model, 
    data = d, 
    estimator = "MLR", 
    missing = "ML"
    )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>%
  print()

# export factor validity
assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std") %$% 
    lambda
```

Shows that PC01_03 doesn't load well. As it's an inverted item that's not surprising. Also from a theoretic perspective it's suboptimal, because it doesn't explicitly focus on privacy, but just concerns in general. Will be deleted.

### CFA 2

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
"
fit <-
  lavaan::sem(
    model = model,
    data = d,
    estimator = "MLR",
    missing = "ML"
    )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>% 
  print()

assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std"
  ) %$% 
  lambda
```

Updated version shows good fit.

## Gratifications general
### Items

Using the participation platform ...

 1.	... had many benefits for me.
 2.	... has paid off for me.
 3.	... was worthwhile.
 4.	... was fun.
 5.	... has brought me further regarding content.

### Distributions

```{r fig.height = 3}
plot_dist(
  item = "GR02"
)

get_desc(
  item = "GR02",
  name = "gratsgen"
)
```

### CFA

```{r}
name <- "gratsgen"
model <- "
  grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
"
fit <- 
  lavaan::sem(
    model = model, 
    data = d, 
    estimator = "MLR", 
    missing = "ML"
    )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>%
  print()

assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std"
  ) %$% 
  lambda
```

## Gratifications specific
### Items

Using the participation platform it has been possible for me ...

_Information_

 1.	... to learn things I would not otherwise have noticed.
 2.	... to hear the opinion of others.
 3.	... to learn how other people tick.

_Relevance_

 4.	... to react to a subject that is very dear to me.
 5.	... to react to a subject that is important to me.
 6.	... to react to a subject that I am affected by.

_Political participation_

 7.	... to engage politically.
 8.	... to discuss political issues.
 9.	... to pursue my political interest.

_Idealism_

 10. ... to try to improve society.
 11. ... to advocate to something meaningful.
 12. ... to serve a good purpose.

_Extrinsic benefits_

 13. ... to do the responsible persons a favor.
 14. ... to soothe my guilty consciences.
 15. ... to fulfil my civic duty.

### Distributions

```{r}
plot_dist(
  item = "GR01",
  n_rows = 2
)

get_desc(
  item = "GR01",
  name = "gratsspec"
)
```

### CFA

```{r}
name <- "gratsspec"
model <- "
  grats_inf =~ GR01_01 + GR01_02 + GR01_03 
  grats_rel =~ GR01_04 + GR01_05 + GR01_06 
  grats_par =~ GR01_07 + GR01_08 + GR01_09
  grats_ide =~ GR01_10 + GR01_11 + GR01_12 
  grats_ext =~ GR01_13 + GR01_14 + GR01_15
  grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext
"
fit <- lavaan::sem(
  model = model, 
  data = d, 
  estimator = "MLR", 
  missing = "ML"
  )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>%
  print()

assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std"
  ) %$% 
  lambda
```

## Privacy deliberation
### Items

Using the participation platform ...

 1. ... I have considered whether I could be disadvantaged by writing a comment.
 2. ... I have considered whether I could be advantaged by writing a comment.
 3. ... I have weighed up the advantages and disadvantages of writing a comment.
 4. ... I have thought about consequences of a possible comment.
 5. ... I have considered whether I should write a comment or not.

### Distributions

```{r fig.height = 3}
plot_dist(
  item = "PD01"
)

get_desc(
  item = "PD01",
  name = "pridel"
)
```

### CFA

```{r}
name <- "pridel"
model <- "
  pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
"
fit <- lavaan::sem(
  model = model, 
  data = d, 
  estimator = "MLR", 
  missing = "ML"
  )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>%
  print()

assign(
  paste0(
    name, 
    "_facval"), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std"
  ) %$%
  lambda
```

## Trust general
### Items

&nbsp;&nbsp;&nbsp;1. The other users seemed trustworthy.  
&nbsp;&nbsp;&nbsp;2. The operators of the participation platform seemed trustworthy.  
&nbsp;&nbsp;&nbsp;3. The website seemed trustworthy.

### Distributions

```{r fig.height = 3}
plot_dist(
  item = "TR02"
)
get_desc(
  item = "TR02",
  name = "trustgen"
)
```

### CFA

```{r}
name <- "trustgen"
model <- "
trust =~ TR02_01 + a*TR02_02 + a*TR02_03
"
fit <- 
  lavaan::sem(
    model = model, 
    data = d, 
    estimator = "MLR", 
    missing = "ML"
    )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>%
  print()

assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std"
  ) %$%
  lambda
```

Note that we constrained Items 5 and Item 9 to be equal. Explanation: First, they are theoretically related. Second, not constraining would yield to just-identified model, for which model fit cannot be interpreted meaningfully.

## Trust specific
### Items

_Community_

 1.	The comments of other users were useful.
 2.	The other users had good intentions.
 3.	I could rely on the statements of other users.

_Provider_

 4.	The operators of the participation platform have done a good job.
 5.	It was important to the operators that the users are satisfied with the participation platform.
 6.	I could rely on the statements of the operators of the participation platform.

_Information System_

 7.	The website worked well.
 8.	I had the impression that my data was necessary for the use of the website.
 9.	I found the website useful.

### Distributions

```{r}
plot_dist(
  item = "TR01",
  n_rows = 2
)
get_desc(
  item = "TR01",
  name = "trustspec"
)
```

### CFA

```{r}
name <- "trust_spec"
model <- "
  trust_community =~ TR01_01 + TR01_02 + TR01_03
  trust_provider =~ TR01_04 + TR01_05 + TR01_06
  trust_system =~ TR01_07 + TR01_08 + TR01_09

  trust =~ trust_community + trust_provider + trust_system
"
fit <- lavaan::sem(
  model = model, 
  data = d, 
  estimator = "MLR", 
  missing = "ML"
  )
```

Model fit:

```{r warning=T}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>% 
  print()

assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std") %$%
  lambda
```

### CFA 2

Because there was Heywoodcase, we now also run a model in which the subdimensions `provider` and `system` are combined.

```{r}
name <- "trustspec"
model <- "
  trust_community =~ TR01_01 + TR01_02 + TR01_03
  trust_provider =~ TR01_04 + TR01_05 + TR01_06 + TR01_07 + TR01_08 + TR01_09
  trust_spec =~ b*trust_community + b*trust_provider
"
fit <- lavaan::sem(
  model = model, 
  data = d, 
  estimator = "MLR", 
  missing = "ML"
  )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>%
  print()

assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std"
  ) %$%
  lambda
```
Warning disappears, results show adequate fit.

## Self-efficacy
### Items

 1. In principle, I felt able to write a comment.
 2. I felt technically competent enough to write a comment.
 3. In terms of the topic, I felt competent enough to express my opinion.
 4. I found it easy to express my opinion regarding the topic.
 5. I found it complicated to write a comment. (reversed)
 6. I was overburdened to write a comment. (reversed)

### Distributions

```{r fig.height = 3}
plot_dist(
  item = "SE01"
)
get_desc(
  item = "SE01",
  name = "selfeff"
)
```

### CFA

```{r}
name <- "self-eff"
model <- "
  self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04 + SE01_05 + SE01_06
"
fit <- 
  lavaan::sem(
    model = model, 
    data = d, 
    estimator = "MLR", 
    missing = "ML"
    )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>%
  print()

assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit,
  what = "std") %$%
  lambda
```

Shows significant misfit. We will delete inverted items, while allowing covariations between Items 1 and 2 (tech-oriented) and Items 3 and 4 (topic-oriented).

### CFA 2

```{r}
name <- "selfeff"
model <- "
  self_eff_pos =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
"
fit <- 
  lavaan::sem(
    model = model, 
    data = d, 
    estimator = "MLR", 
    missing = "ML"
    )
```

Model fit:

```{r}
facval <- 
  fit_tab(
    fit, 
    reliability = TRUE, 
    scaled = TRUE
    ) %T>% 
  print()

assign(
  paste0(
    name, 
    "_facval"
    ), 
  facval
  )
```

Factor loadings:

```{r}
inspect(
  fit, 
  what = "std") %$%
  lambda
```

Adapted version shows better and adequate fit.

## Communication
### Density plot

```{r fig.height = 3}
ggplot(
  gather(
    select(
      d, 
      words
      )
    ), 
  mapping = aes(
    x = value
    )
  ) +
  geom_density() +
  facet_wrap(
    ~key, 
    nrow = 1
    ) + 
  theme_bw()
```

We see that Communication is severely skewed. Let's calculate how many participants actually commented.

```{r}
no_words_perc <- 
  d %>% 
  select(words) %>% 
  table() %>% 
  prop.table() %>% 
  .[1] %>% 
  unlist()

words_m <- 
  mean(
    d$words, 
    na.rm = TRUE
    )
```

Will hence be log-scaled for SEMs.

## Communication Logged

```{r}
ggplot(
  gather(
    select(
      d, 
      words_log
      )
    ), 
  mapping = aes(
    x = value
    )
  ) +
  geom_density() +
  facet_wrap(
    ~key, 
    nrow = 1
    ) + 
  theme_bw()
```

## Baseline model

In what follows, please find the results of all variables combined in one model. This model will be used to extract factor scores.

```{r cache=F}
model_baseline <- "
  pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
  grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
  grats_inf =~ GR01_01 + GR01_02 + GR01_03 
  grats_rel =~ GR01_04 + GR01_05 + GR01_06 
  grats_par =~ GR01_07 + GR01_08 + GR01_09
  grats_ide =~ GR01_10 + GR01_11 + GR01_12 
  grats_ext =~ GR01_13 + GR01_14 + GR01_15
  grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext
  pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
  trust_gen =~ TR02_01 + TR02_02 + TR02_03
  trust_community =~ TR01_01 + TR01_02 + TR01_03
  trust_provider =~ TR01_04 + TR01_05 + TR01_06 + TR01_07 + TR01_08 + TR01_09
  trust_spec =~ b*trust_community + b*trust_provider
  self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
    SE01_01 ~~ x*SE01_02
    SE01_03 ~~ x*SE01_04
  Words_log =~ words_log

  Words_log ~~ a1*pri_con + b1*grats_gen + c1*pri_delib + d1*self_eff + e1*trust_spec + f1*trust_gen + g1*grats_spec
"

fit_baseline <- 
  lavaan::sem(
    model_baseline, 
    data = d, 
    missing = "ML"
    )

summary(
  fit_baseline, 
  standardized = TRUE, 
  fit.measures = TRUE
  )
```

# Descriptive analyses

We first report the factor validity of all variables combined.

```{r cache=F}
# extract model factor scores / predicted values for items & calc means
d_fs <- 
  lavPredict(
    fit_baseline, 
    type = "ov"
    ) %>% 
  as.data.frame() %>% 
  mutate(
    version = d$version, 
    grats_gen_fs = rowMeans(select(., starts_with("GR02"))),
    grats_spec_fs = rowMeans(select(., starts_with("GR01"))), 
    pri_con_fs = rowMeans(select(., starts_with("PC01"))),
    trust_gen_fs = rowMeans(select(., starts_with("TR02"))),
    trust_spec_fs = rowMeans(select(., starts_with("TR01"))),
    pri_del_fs = rowMeans(select(., starts_with("PD01"))),
    self_eff_fs = rowMeans(select(., starts_with("SE01")))) %>%
  select(
    version, 
    pri_con_fs, 
    grats_gen_fs, 
    grats_spec_fs, 
    pri_del_fs, 
    self_eff_fs, 
    trust_gen_fs, 
    trust_spec_fs, 
    words_log
    )

# combine d with d factor scores
d %<>% 
  cbind(
    select(
      d_fs, 
      -version, 
      -words_log
      )
    )

# add centered predictors for interaction analyses later
d %<>%
  mutate(
    pri_con_fs_c = scale(pri_con_fs, scale = FALSE),
    grats_spec_fs_c = scale(grats_spec_fs, scale = FALSE),
    pri_del_fs_c = scale(pri_del_fs, scale = FALSE),
    trust_gen_fs_c = scale(trust_gen_fs, scale = FALSE),
    self_eff_fs_c = scale(self_eff_fs, scale = FALSE),
    male_c = scale(male, scale = FALSE),
    edu_c = scale(edu, scale = FALSE),
    age_c = scale(age, scale = FALSE),
    con_x_del = pri_con_fs_c * pri_del_fs_c,
    grats_x_del = grats_spec_fs_c * pri_del_fs_c
    )

# rename for plotting
d_fs %<>% 
  set_names(
    c(
      "version", 
      var_names_breaks
      )
    )

# means of model predicted values
des <- 
  rbind(
    des_pricon,
    des_gratsgen,
    des_gratsspec,
    des_pridel,
    des_trustgen,
    des_trustspec,
    des_selfeff
    )

facval_tab <- 
  rbind(
    pricon_facval,
    gratsgen_facval,
    gratsspec_facval,
    pridel_facval,
    selfeff_facval,
    trustgen_facval,
    trustspec_facval
    ) %$%
  cbind(
    des[-c(8), ],
    .
    ) %>%
  set_rownames(
    var_names[-c(8)]
    )

facval_tab %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%")
```

In what follows, we report zero-order correlations, distributions, and scatterplots of the variables' factor scores.

```{r fig.height=9, fig.width=9}
corr_plot <- 
  ggpairs(
    select(d_fs, -version),
    upper = list(
      continuous = cor_plot
      ),
    lower = list(
      continuous = wrap(
        td::scat_plot, 
        coords = c(1, 7, 0, 7)
        )
      )
    ) + 
  theme_bw()
print(corr_plot)
ggsave("figures/results/cor_plot.png")
```

# Power analyses

In what follows, we report power analyses for our study. Please note that we conduct a rudimentary power-analysis, assuming bivariate correlations. To explain, at the time we were not yet aware of the existence of power analyses for multivariate structural equation models.

```{r include=F}
# define pwr-criteria
z_crit <- 1.96
alpha <- .05
power_desired <- .95
r_sesoi <- .10
```

We first estimate the sample size necessary to find small effects in 95% of all cases.

```{r}
# estimate pwr-samplesize
pwr.r.test(
  r = r_sesoi, 
  sig.level = alpha, 
  power = power_desired, 
  alternative = "greater"
  ) %T>%
  print %$% 
  n %>% 
  round(0) ->
  n_desired
```

We then compute the power we have achieved with our finale sample size to detect small effects.

```{r}
# compute pwr-achieved
pwr.r.test(
  n = n_final, 
  r = r_sesoi, 
  sig.level = alpha, 
  alternative = "greater"
  ) %T>%
  print %$% 
  power %>% 
  round(2) ->
  power_achieved
```

We finally compute what effect size we are likely to find in 95% of all cases given our final sample size.

```{r}
# estimate pwr-sensitivity
pwr.r.test(
  n = n_final, 
  power = power_desired, 
  sig.level = alpha, 
  alternative = "greater"
  ) %T>%
  print %$% 
  r %>% 
  round(2) ->
  r_sensitive
```

# Assumptions
## Multivariate normal distribution

```{r}
# create subset of data with all items that were used
items_used <- 
  c(
    "words", 
    "GR02_01", "GR02_02", "GR02_03", "GR02_04", "GR02_05", 
    "PC01_01", "PC01_02", "PC01_04", "PC01_05", "PC01_06", "PC01_07", 
    "TR01_01", "TR01_02", "TR01_03", "TR01_04", "TR01_05", "TR01_06", "TR01_07", "TR01_08", "TR01_09",
    "TR02_01", "TR02_02", "TR02_03", 
    "PD01_01", "PD01_02", "PD01_03", "PD01_04", "PD01_05", 
    "SE01_01", "SE01_02", "SE01_03", "SE01_04", 
    "male", "age", "edu"
    )
d_sub <- d[, items_used]

# test multivariate normal distribution
mvn_result <- mvn(d_sub, mvnTest = "mardia")
mvn_result$multivariateNormality
```

Shows that multivariate normal distribution is violated. We hence use maximum likelihood estimation with robust standard errors and a Satorra-Bentler scaled test statistic.

## Influential cases

[Note: These lines stopped working after some time, potentially due to changes in package.]

In what follows we test for influential cases in the baseline model, to detect potentially corrupt data (e.g., people who provided response patterns). Specifically, we compute Cook's distance.

```{r eval=F}
cooks_dis <- 
  gCD(
    d, 
    model_baseline
    ) %T>% 
  plot()
```

The following ten cases have a particularly strong influence on the baseline model.

```{r eval=F}
infl_cases <- 
  invisible(
    rownames(
      print(
        cooks_dis
        )
      )
    )
```

Let's inspect these cases.

```{r eval=F}
infl_cases_tokens <- 
  d[infl_cases, "case_token"] %>% 
  as_vector()

d %>% 
  filter(
    case_token %in% infl_cases_tokens
    ) %>% 
  select(
    case_token, 
    GR01_01:SE01_06, 
    topics_entered:reactions, 
    -SO01_01, 
    TIME_SUM_t1, 
    TIME_SUM_t2
    ) %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%")
```

These data do not reveal potential cases of response patterns. Indeed, answer times suggest that respondents were diligent.

# Results
## Preregistered
### Privacy calculus

```{r}
model <- "
  pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 
  grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
  pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
  self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
  trust_community =~ TR01_01 + TR01_02 + TR01_03
  trust_provider =~ TR01_04 + TR01_05 + TR01_06 + TR01_07 + TR01_08 + TR01_09
  trust_spec =~ b*trust_community + b*trust_provider

words_log ~ a1*pri_con + b1*grats_gen + c1*pri_delib + d1*self_eff + e1*trust_spec

# Covariates
words_log + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_01 + TR01_02 + TR01_03 + TR01_04 + TR01_05 + TR01_06 + TR01_07 + TR01_08 + TR01_09 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu

# Covariances
male ~~ age + edu
age ~~ edu
"

fit_prereg <- 
  lavaan::sem(
    model, 
    data = d, 
    estimator = "MLR", 
    missing = "ML"
    )

summary(
  fit_prereg, 
  fit = TRUE, 
  std = TRUE
  )

rsquare_fit_prereg <- 
  inspect(
    fit_prereg, 
    what = "rsquare"
    )["words"]
```

Results show that there's only one significant predictor of Communication, being self-efficacy. The other predictors are in the direction as planned, albeit not significant. Trust, however, shows the inverse relation as effect, that is, more trust, less communication.

### Effects of popularity cues
#### Confidence intervals

The easiest way to assess the effect of the experimental manipulation on the variables is by visualizing their means. If 83% confidence intervals don't overlap, the variables differ significantly across the conditions. One can quickly see that there aren't any major effects.

```{r}
# violin plot
fig_fs_m <- 
  ggplot(
    gather(
      d_fs, 
      variable, 
      value, 
      -version) %>% 
      mutate(
        variable = factor(
          variable, 
          levels = var_names_breaks
          )
        ),
    aes(
      x = version, 
      y = value, 
      fill = version
      )
    ) +
  geom_violin(
    trim = TRUE
    ) +
  stat_summary(
    fun.y = mean, 
    geom = "point"
    ) +
  stat_summary(
    fun.data = mean_se, 
    fun.args = list(mult = 1.39), 
    geom = "errorbar", 
    width = .5
    ) + 
  facet_wrap(
    ~ variable, 
    nrow = 2
    ) + 
  theme_bw() +
  theme(
    axis.title.y = element_blank(),
    axis.title.x = element_blank(),
    plot.title = element_text(hjust = .5),
    panel.spacing = unit(.9, "lines"),
    text = element_text(size = 12),
    legend.position="none",
    legend.title = element_blank()) +
  coord_cartesian(
    ylim = c(0, 7)
    ) +
  scale_fill_brewer(
    palette = "Greys"
    )
ggsave(
  "figures/results/violin_plot.png", 
  width = 8, 
  height = 6
  )
fig_fs_m
```

#### SEM

In what follows, we also report explicit statistical tests of the differences between the conditions using contrasts.

**Like & Like-Dislike vs. Control**

```{r}
model <- "
  pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
  grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
  pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
  self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
  trust_community =~ TR01_01 + TR01_02 + TR01_03
  trust_provider =~ TR01_04 + TR01_05 + TR01_06 + TR01_07 + TR01_08 + TR01_09
  trust_spec =~ b*trust_community + b*trust_provider

  pri_con + grats_gen + pri_delib + self_eff + trust_spec ~ like + likedislike
  words ~ a*pri_con + b*grats_gen + c*pri_delib + d*self_eff + e*trust_spec + f*like + g*likedislike

# Covariates
words + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_01 + TR01_02 + TR01_03 + TR01_04 + TR01_05 + TR01_06 + TR01_07 + TR01_08 + TR01_09 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"

fit_lik_ctrl <- 
  lavaan::sem(
    model = model, 
    data = d, 
    estimator = "MLR", 
    missing = "ML", 
    fixed.x = FALSE
    )

summary(
  fit_lik_ctrl, 
  fit = TRUE, 
  std = TRUE
  )
```

No significant effects of popularity cues on privacy calculus.

**Like-Dislike & Control vs. Like**

```{r}
model <- "
  pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
  grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
  pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
  self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
  trust_community =~ TR01_01 + TR01_02 + TR01_03
  trust_provider =~ TR01_04 + TR01_05 + TR01_06 + TR01_07 + TR01_08 + TR01_09
  trust_spec =~ b*trust_community + b*trust_provider

  pri_con + grats_gen + pri_delib + self_eff + trust_spec + words ~ likedislike + control

# Covariates
words + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + TR01_01 + TR01_02 + TR01_03 + TR01_04 + TR01_05 + TR01_06 + TR01_07 + TR01_08 + TR01_09 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 ~ male + age + edu
"
fit_lik_ctrl <- 
  lavaan::sem(
    model = model, 
    data = d, 
    estimator = "MLR", 
    missing = "ML")

summary(
  fit_lik_ctrl, 
  fit = TRUE, 
  std = TRUE
  )
```

No significant effects of popularity cues on privacy calculus.

## Exploratory analyses

The bivariate relations showed several strong correlations. 
People who trusted the providers more also experienced more gratifications (_r_ = .79).
In multiple regression, we analze the potential effect of one variable on the outcome while holding all others constant. 
In this case, I tested whether trust increases communication while holding constant gratifications, privacy concerns, privacy deliberations, and self-efficacy. 
Given the close theoretical and empirical interrelations between the predictors, this creates an unlikely and artifical scenario. 
If people say loose trust in a provider, they will likely experience more concerns and less benefits. 
But it is not even necessary to control for these mediating factors. 
When trying to analyze the causal effect, it is necessary to control for counfounding variables, but importantly not for mediators [@rohrerThinkingClearlyCorrelations2018].
Confunding variables impact both the independent variable and the outcome. 
Sociodemographic variables are often ideal candidates for confounders.
For example, men are generally less concerned about their privacy and post more online.
Hence, controlling for gender hence helps isolate the actual causal effect. 
As a result, I reran the analyses controlling for confounding variables that are not mediators.

### Experimental Factors
#### Words

We first compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d,
  name = "wrds_ctrl",
  outcome = "words", 
  predictor = "version",
  y_lab = "Words",
  x_lab = "Experimental conditions"
  )
```

Let's find out by how much people in like and like plus dislike condition wrote fewer words.

```{r}
words_m_ctrl <- 
  fig_wrds_ctrl$data %>% 
  filter(effect1__ == "Control") %>% 
  select(estimate__)

words_m_lk <- 
  fig_wrds_ctrl$data %>% 
  filter(effect1__ == "Like") %>% 
  select(estimate__)

words_m_lkdslk <- 
  fig_wrds_ctrl$data %>% 
  filter(effect1__ == "Like & Dislike") %>% 
  select(estimate__)

effect_lk_wrds <- words_m_ctrl - words_m_lk
effect_lkdslk_wrds <- words_m_ctrl - words_m_lkdslk

effect_lk_prcnt <- 
  (effect_lk_wrds / words_m_ctrl) * 100 %>% 
  round()
effect_lkdlk_prcnt <- 
  (effect_lkdslk_wrds / words_m_ctrl) * 100 %>% 
  round()
```

Compared to the control condition, participants wrote `r effect_lkdlk_prcnt` percent fewer words on the platform with likes and dislikes.

We then compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d, 
  name = "wrds_lkdslk",
  outcome = "words", 
  predictor = "version_lkdslk",
  y_lab = "Words",
  x_lab = "Experimental conditions"
  )
```

#### Privacy Concerns

We first compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d,
  name = "pricon_ctrl",
  outcome = "pri_con_fs", 
  predictor = "version",
  y_lab = "Privacy concerns",
  x_lab = "Experimental conditions"
  )
```

We then compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d, 
  name = "pricon_lkdslk",
  outcome = "pri_con_fs", 
  predictor = "version_lkdslk",
  y_lab = "Privacy concerns",
  x_lab = "Experimental conditions"
  )
```

#### Expected Gratifications

We first compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d,
  name = "grats_ctrl",
  outcome = "grats_gen_fs", 
  predictor = "version",
  y_lab = "Gratifications",
  x_lab = "Experimental conditions"
  )
```

We then compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d, 
  name = "grats_lkdslk",
  outcome = "grats_gen_fs", 
  predictor = "version_lkdslk",
  y_lab = "Gratifications",
  x_lab = "Experimental conditions"
  )
```

#### Privacy Deliberation

We first compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d,
  name = "pridel_ctrl",
  outcome = "pri_del_fs", 
  predictor = "version",
  y_lab = "Privacy deliberation",
  x_lab = "Experimental conditions"
  )
```

We then compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d, 
  name = "pridel_lkdslk",
  outcome = "pri_del_fs", 
  predictor = "version_lkdslk",
  y_lab = "Privacy deliberation",
  x_lab = "Experimental conditions"
  )
```

#### Trust

We first compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d,
  name = "trust_ctrl",
  outcome = "trust_gen_fs", 
  predictor = "version",
  y_lab = "Trust",
  x_lab = "Experimental conditions"
  )
```

We then compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d, 
  name = "trust_lkdslk",
  outcome = "trust_gen_fs", 
  predictor = "version_lkdslk",
  y_lab = "Trust",
  x_lab = "Experimental conditions"
  )
```

#### Self-Efficacy

We first compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d,
  name = "selfeff_ctrl",
  outcome = "self_eff_fs", 
  predictor = "version",
  y_lab = "Self-efficacy",
  x_lab = "Experimental conditions"
  )
```

We then compare likes and likes plus dislikes to the control condition.

```{r}
run_hurdles(
  object = d, 
  name = "selfeff_lkdslk",
  outcome = "self_eff_fs", 
  predictor = "version_lkdslk",
  y_lab = "Self-efficacy",
  x_lab = "Experimental conditions"
  )
```

### Words
#### Privacy Concerns

```{r}
run_hurdles(
  object = d, 
  name = "wrds_pricon",
  outcome = "words", 
  predictor = "pri_con_fs",
  y_lab = "Words",
  x_lab = "Privacy concerns"
  )
```

Let's explore  how number of communicated words differ for people very much concerned versus not concerned at all.

```{r}
# let's extract effects for 1
wrds_pricon_1 <- 
  fig_wrds_pricon %$%
  data %>% 
  filter(
    pri_con_fs ==
      min(.$pri_con_fs)
  ) %$%
  estimate__ %>% 
  round(0)

# let's extract effects for 7
wrds_pricon_7 <- 
  fig_wrds_pricon %$%
  data %>% 
  filter(
    pri_con_fs ==
      max(.$pri_con_fs)
  ) %$%
  estimate__ %>% 
  round(0)
```

People who were not concerned at all communicated on average `r wrds_pricon_1 %>% round(0) %>% abs()` words, whereas people strongly concerned communicated on average `r wrds_pricon_7 %>% round(0)` words.

#### Expected Gratifications

```{r}
run_hurdles(
  object = d, 
  name = "wrds_grats",
  outcome = "words", 
  predictor = "grats_gen_fs",
  y_lab = "Words",
  x_lab = "Expected gratifications"
  )
```

#### Privacy Deliberation

```{r}
run_hurdles(
  object = d, 
  name = "wrds_pridel",
  outcome = "words", 
  predictor = "pri_del_fs",
  y_lab = "Words",
  x_lab = "Privacy deliberation"
  )
```

#### Trust

```{r}
run_hurdles(
  object = d, 
  name = "wrds_trust",
  outcome = "words", 
  predictor = "trust_gen_fs",
  y_lab = "Words",
  x_lab = "Trust"
  )
```

#### Self-Efficacy

```{r}
run_hurdles(
  object = d, 
  name = "wrds_selfeff",
  outcome = "words", 
  predictor = "self_eff_fs",
  y_lab = "Words",
  x_lab = "Self-Efficacy"
  )
```

The effects is markedly exponential. Let's inspect the changes across values in self-efficacy.

```{r}
effects_wrds_selfeff <- 
  slopes(
    fit_wrds_selfeff, 
    newdata = datagrid(
      self_eff_fs = c(1, 6, 7)
      ), 
   grid_type = "counterfactual"
    ) %>%
  filter(
    term == "self_eff_fs"
    ) %T>%
  print()

# let's extract effects for 1
effects_wrds_selfeff_12 <- 
  effects_wrds_selfeff %>% 
  filter(
    self_eff_fs == 1 &
      term == "self_eff_fs") %>% 
  select(estimate)

# let's extract effects for 7
effects_wrds_selfeff_67 <- 
  effects_wrds_selfeff %>% 
  filter(
    self_eff_fs == 6 &
      term == "self_eff_fs") %>% 
  select(estimate)

```

Whereas a change in self-efficacy from 1 to 2 led to an increase of `r effects_wrds_selfeff_12 %>% round(0)` words, a change from 6 to 7 led to an increase of `r effects_wrds_selfeff_67 %>% round(0)` words.

Let's explore  how number of communicated words differ for people very much self-efficacious versus not self-efficacious at all.

```{r}
# let's extract effects for 1
wrds_selfeff_1 <- 
  fig_wrds_selfeff %$%
  data %>% 
  filter(
    self_eff_fs ==
      min(.$self_eff_fs)
  ) %$%
  estimate__ %>% 
  round(0)

# let's extract effects for 7
wrds_selfeff_7 <- 
  fig_wrds_selfeff %$%
  data %>% 
  filter(
    self_eff_fs ==
      max(.$self_eff_fs)
  ) %$%
  estimate__ %>% 
  round(0)
```

People who reported no self-efficacy at all communicated on average `r wrds_selfeff_1 %>% round(0) %>% abs()` words, whereas people with very high self-efficacy communicated on average `r wrds_selfeff_7 %>% round(0)` words.

## Visualization

```{r}
fig_results <- 
  plot_grid(
    fig_wrds_pricon, 
    fig_wrds_grats, 
    fig_wrds_pridel, 
    fig_wrds_trust, 
    fig_wrds_selfeff,
    fig_wrds_ctrl
    ) %T>% 
  print()

ggsave("figures/results/effects.pdf")
ggsave("figures/results/effects.png")
```

## Table

Make table of slopes

```{r}
tab_slopes <- 
  rbind(
    slopes_wrds_pricon,
    slopes_wrds_grats,
    slopes_wrds_pridel,
    slopes_wrds_trust,
    slopes_wrds_selfeff,
    slopes_wrds_ctrl,
    slopes_wrds_lkdslk,
    slopes_pricon_ctrl,
    slopes_pricon_lkdslk,
    slopes_grats_ctrl,
    slopes_grats_lkdslk,
    slopes_pridel_ctrl,
    slopes_pridel_lkdslk,
    slopes_trust_ctrl,
    slopes_trust_lkdslk,
    slopes_selfeff_ctrl,
    slopes_selfeff_lkdslk
    ) %>% 
  as.data.frame() %>% 
  mutate(
    term = ifelse(
      .$contrast == "Like - Control", 
      "Like - Control",
      ifelse(
        .$contrast == "Like & Dislike - Control", 
        "Like & Dislike - Control",
        ifelse(
          .$contrast == "Like - Like & Dislike", 
          "Like - Like & Dislike", 
          .$term
          )
        )
      ),
    Predictor = recode(
      term, 
      "Like - Control" = "Like vs.control",
      "Like & Dislike - Control" = "Like & dislike vs.control",
      "Like - Like & Dislike" = "Like vs.like & dislike",
      "pri_con_fs" = "Privacy concerns",
      "pri_del_fs" = "Privacy deliberation",
      "grats_gen_fs" = "Expected gratifications",
      "trust_gen_fs" = "Trust",
      "self_eff_fs" = "Self-efficacy"
    ),
    Outcome = recode(
      Outcome,
      "pri_con_fs" = "Privacy concerns",
      "pri_del_fs" = "Privacy deliberation",
      "grats_gen_fs" = "Expected gratifications",
      "trust_gen_fs" = "Trust",
      "self_eff_fs" = "Self-efficacy"
    )
  ) %>% 
  filter(
    Predictor %in% c(
      "Like vs.control",
      "Like & dislike vs.control",
      "Like vs.like & dislike",
      "Privacy concerns",
      "Privacy deliberation",
      "Expected gratifications",
      "Trust",
      "Self-efficacy")
  ) %>% 
  mutate(
    Predictor = factor(Predictor, levels = c(
      "Self-efficacy",
      "Trust",
      "Privacy deliberation",
      "Expected gratifications",
      "Privacy concerns",
      "Like vs.like & dislike",
      "Like & dislike vs.control",
      "Like vs.control"
      )
    ),
    Estimate = 
      ifelse(
        Outcome == "words",
        round(estimate),
        estimate
      ),
    LL = 
      ifelse(
        Outcome == "words",
        round(conf.low),
        conf.low
      ),
    UL = 
      ifelse(
        Outcome == "words",
        round(conf.high),
        conf.high
      ),
  ) %>% 
  select(
    Outcome,
    Predictor, 
    Estimate,
    LL, 
    UL
    )

tab_slopes %>% 
  kable() %>% 
  kable_styling("striped") %>% 
  scroll_box(width = "100%")
```

```{r SaveData, include=F, cache=F}
write.csv(
  d, 
  file = "data/data.csv", 
  row.names = FALSE
  )

save.image(
  file = "data/workspace_full.RData"
)

save(
  d_raw, d_all, d_fs, d
  , fit_baseline, fit_prereg
  # , fig_fs_m, fig_fs_cor, fig_coeffs
  , des_pridel
  , facval_tab
  , n_comments, n_desired, n_posts, n_speeding, n_time, n_words
  , n_users, n_final
  # , infl_cases, infl_cases_tokens
  , resp_pattern, n_resp_pattern
  , n_t1, age_t1_m, male_t1_m, college_t1_m
  , n_t2, age_t2_m, male_t2_m, college_t2_m
  , n_matched, n_final, age_final_m, male_final_m, college_final_m, words_m
  , r_sesoi, r_sensitive, power_achieved
  , no_words_perc
  , effects_wrds_selfeff_12, effects_wrds_selfeff_67, effect_lkdlk_prcnt
  , wrds_pricon_7, wrds_pricon_1, wrds_selfeff_1, wrds_selfeff_7
  , tab_slopes
  , file = "data/workspace.RData"
)
```