### 2023-04-11 analyses_explorations.Rmd

 Explorations

### Model "Adapted"

Building on the preregistered model, instead of general gratifications and specific trust, we now use specific gratifications and general trust.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_gen =~ TR01_01 + TR01_05 + TR01_09

COMM_log ~ a1*pri_con + b1*grats_spec + c1*pri_delib + d1*self_eff + e1*trust_gen

# Covariates
COMM_log  + GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 + TR01_01 + TR01_05 + TR01_09 ~ male + age + edu

# Covariances
male ~~ age + edu
age ~~ edu
"
fit_adapted <- lavaan::sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit_adapted, fit = TRUE, std = TRUE)
rsquare_fit_adapted <- inspect(fit_adapted, what = "rsquare")["comm"]
```

Effects become a bit larger and significant, which seems to make sense. However, this also includes trust, which is a significant negative predictor of communication.

### Model "Adapted 2"

Building on the preregistered model, instead of general gratifications and specific trust, we now use specific gratifications and general trust.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07
grats_gen =~ GR02_01 + GR02_02 + GR02_03 + GR02_04 + GR02_05
pri_delib =~ PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_gen =~ TR01_01 + TR01_05 + TR01_09

COMM_log ~ a1*pri_con + b1*grats_gen + c1*pri_delib + d1*self_eff + e1*trust_gen

# Covariates
COMM_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 + PD01_01 + PD01_02 + PD01_03 + PD01_04 + PD01_05 + SE01_01 + SE01_02 + SE01_03 + SE01_04 + TR01_01 + TR01_05 + TR01_09 ~ male + age + edu

# Covariances
male ~~ age + edu
age ~~ edu
"
fit_adapted_2 <- lavaan::sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit_adapted_2, fit = TRUE, std = TRUE)
rsquare_fit_adapted_2 <- inspect(fit_adapted_2, what = "rsquare")["comm"]
```

### Model "Simple"

We now use only two variables, namely specific gratifications and privacy concerns.

```{r}
model <- "
pri_con =~ PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07  
grats_inf =~ GR01_01 + GR01_02 + GR01_03 
grats_rel =~ GR01_04 + GR01_05 + GR01_06 
grats_par =~ GR01_07 + GR01_08 + GR01_09
grats_ide =~ GR01_10 + GR01_11 + GR01_12 
grats_ext =~ GR01_13 + GR01_14 + GR01_15
grats_spec =~ grats_inf + grats_rel + grats_par + grats_ide + grats_ext

COMM_log ~ a1*pri_con + b1*grats_spec

# Covariates
COMM_log + GR01_01 + GR01_02 + GR01_03 + GR01_04 + GR01_05 + GR01_06 + GR01_07 + GR01_08 + GR01_09 + GR01_10 + GR01_11 + GR01_12 + GR01_13 + GR01_14 + GR01_15 + PC01_01 + PC01_02 + PC01_04 + PC01_05 + PC01_06 + PC01_07 ~ male + age + edu

# Covariances
male ~~ age + edu
age ~~ edu
"
fit_simple <- lavaan::sem(model, data = d, estimator = "MLR", missing = "ML")
summary(fit_simple, fit = TRUE, std = TRUE)
rsquare_fit_simple <- inspect(fit_simple, what = "rsquare")["comm"]
```

Works as hypothesized.

### Model "Interaction complex"

We now estimate an interaction model, as suggested by an anonymous reviewer. To estimate the interaction we'll use factor scores.

```{r}
model <- "
self_eff =~ SE01_01 + SE01_02 + SE01_03 + SE01_04
  SE01_01 ~~ x*SE01_02
  SE01_03 ~~ x*SE01_04
trust_gen =~ TR01_01 + TR01_05 + TR01_09

COMM_log ~ a1*pri_con_fs_c + b1*grats_spec_fs_c + c1*pri_del_fs_c + con_x_del + grats_x_del + d1*self_eff + e1*trust_gen

# Covariates
COMM_log + SE01_01 + SE01_02 + SE01_03 + SE01_04 + TR01_01 + TR01_05 + TR01_09 + pri_con_fs_c + grats_spec_fs_c + pri_del_fs_c + con_x_del ~ male + age + edu

# Covariances
pri_con_fs_c ~~ grats_spec_fs_c + pri_del_fs_c + con_x_del + grats_x_del + self_eff + trust_gen
grats_spec_fs_c ~~ pri_del_fs_c + con_x_del + grats_x_del + self_eff + trust_gen
pri_del_fs_c ~~ con_x_del + grats_x_del + self_eff + trust_gen
con_x_del ~~ grats_x_del + self_eff + trust_gen
grats_x_del ~~ self_eff + trust_gen
self_eff ~~ trust_gen
male ~~ age + edu
age ~~ edu
"

fit_inter <- lavaan::sem(model, data = d, estimator = "MLR", missing = "ML", missing = "ML")
summary(fit_inter, fit = TRUE, std = TRUE)
rsquare_fit_inter <- inspect(fit_inter, what = "rsquare")["comm"]
```

Results are interesting. Both interaction effects are significant. If Deliberation increases, the effect of concerns on communication increase (net effect). If Deliberation increases, the effect of gratifications on communication decrease. Hence, could imply that if people think more, concerns more relevant than benefits.

### Model "Interaction simple"

We now estimate an interaction model, as suggested by an anonymous reviewer. To estimate the interaction we'll use factor scores.

```{r}
model <- "
COMM_log ~ a1*pri_con_fs_c + b1*grats_spec_fs_c + c1*pri_del_fs_c + con_x_del + grats_x_del

# Covariates
COMM_log + pri_con_fs_c + grats_spec_fs_c + pri_del_fs_c + con_x_del ~ male + age + edu

# Covariances
pri_con_fs_c ~~ grats_spec_fs_c + pri_del_fs_c + con_x_del + grats_x_del
grats_spec_fs_c ~~ pri_del_fs_c + con_x_del + grats_x_del
pri_del_fs_c ~~ con_x_del + grats_x_del
con_x_del ~~ grats_x_del
male ~~ age + edu
age ~~ edu
"

fit_inter_simple <- lavaan::sem(model, data = d, estimator = "MLR", missing = "ML", missing = "ML")
summary(fit_inter_simple, fit = TRUE, std = TRUE)
rsquare_fit_inter_simple <- inspect(fit_inter_simple, what = "rsquare")["comm"]
```

Results are interesting. Both interaction effects are significant. If Deliberation increases, the effect of concerns on communication increase (net effect). If Deliberation increases, the effect of gratifications on communication decrease. Hence, could imply that if people think more, concerns more relevant than benefits.

### Model Comparisons

```{r}
model_comp_tab <- data.frame(
  preregistered = fitmeasures(fit_prereg, c("BIC", "AIC")),
  adapted = fitmeasures(fit_adapted, c("BIC", "AIC")),
  simple = fitmeasures(fit_simple, c("BIC", "AIC"))) %>% 
  t() %T>%
  print()
```

### Model "BRMS"

Because the dependent variable is a zero-inflated Poisson distribution, we'll also run Bayesian models, which can explicitly account for this distribution.

#### Adapted Model

```{r}
fit_brms <- brm(COMM ~ pri_con_fs + grats_gen_fs + pri_del_fs + self_eff_fs + trust_gen_fs + age + male + edu, data = d, family = zero_inflated_poisson())
d_predicted <- predict(fit_brms) %>% as.data.frame()
summary(fit_brms)
```

Shows very similar results compared to SEM. Again, trust a negative predictor.

Let's also look at how values are predicted to see how model performs.

```{r}
d_predicted <- predict(fit_brms) %>% as.data.frame()
ggplot(d_predicted, aes(Estimate)) + 
  geom_density()
```

Let's also plot actual values.

```{r}
# plot actual values
ggplot(d, aes(COMM)) + 
  geom_density()
```

Model-predicted values don't closely match actual values.

#### Without Trust

Because trust has a negative coefficient, which seems generally implausible, we also run a model without trust.

```{r}
fit_brms <- 
  brm(
    COMM ~ pri_con_fs + grats_spec_fs + pri_del_fs + self_eff_fs + age + male + edu, 
    data = d, 
    family = zero_inflated_poisson()
    )
d_predicted <- 
  predict(fit_brms) %>% 
  as.data.frame()
summary(fit_brms)
```

Also doesn't make much sense, privacy concern now positive predictor. Self-Efficacy a strong predictor of behavior.

#### Simple model

```{r}
fit_brms_simple <- brm(COMM ~ pri_con_fs + grats_spec_fs + age + male + edu, data = d, family = zero_inflated_poisson())
summary(fit_brms_simple)
```

Also works, showing somewhat stronger effects.

#### Preregistered

```{r}
fit_brms_hrdl_prereg <- 
  brm(
    bf(
      COMM ~ 
        1 + 
        pri_con_fs + 
        grats_spec_fs + 
        pri_del_fs + 
        trust_gen_fs +
        self_eff_fs +
        age + 
        male + 
        edu, 
      hu ~ 
        1 + 
        pri_con_fs + 
        grats_spec_fs + 
        pri_del_fs + 
        trust_gen_fs +
        self_eff_fs +
        age + 
        male + 
        edu
    ), 
    data = d, 
    family = hurdle_gamma()
  )

```

```{r}
plot(fit_brms_hrdl_prereg, ask = FALSE)
```

```{r}
model_diagnostics(fit_brms_hrdl_prereg)
```

```{r}
summary(fit_brms_hrdl_prereg)
```

```{r}
conditional_effects(fit_brms_hrdl_prereg, ask = FALSE)
```

#### Adapted

```{r}
fit_brms_hrdl_adapted <- 
  brm(
    bf(
      COMM ~ 
        1 + 
        pri_con_fs + 
        grats_spec_fs + 
        pri_del_fs + 
        self_eff_fs +
        age + 
        male + 
        edu, 
      hu ~ 
        1 + 
        pri_con_fs + 
        grats_spec_fs + 
        pri_del_fs + 
        self_eff_fs +
        age + 
        male + 
        edu
    ), 
    data = d, 
    family = hurdle_gamma()
  )

```

```{r}
plot(fit_brms_hrdl_adapted, ask = FALSE)
```

```{r}
model_diagnostics(fit_brms_hrdl_adapted)
```

```{r}
summary(fit_brms_hrdl_adapted)
```

```{r}
conditional_effects(fit_brms_hrdl_adapted, ask = FALSE)
```

#### 3-way Interaction

```{r}
fit_brms_hrdl_inter <- 
  brm(
    bf(
      COMM ~ 
        1 + 
        pri_con_fs_c * 
        grats_spec_fs_c * 
        pri_del_fs_c + 
        self_eff_fs_c +
        age_c + 
        male_c + 
        edu_c, 
      hu ~ 
        1 + 
        pri_con_fs_c * 
        grats_spec_fs_c * 
        pri_del_fs_c + 
        self_eff_fs_c +
        age_c + 
        male_c + 
        edu_c
    ), 
    data = d, 
    family = hurdle_gamma()
  )

```

```{r}
plot(fit_brms_hrdl_inter, ask = FALSE)
```

```{r}
model_diagnostics(fit_brms_hrdl_inter)
```

```{r}
summary(fit_brms_hrdl_inter)
```

```{r}
conditional_effects(fit_brms_hrdl_inter, ask = FALSE)
```

#### Simple hurdle model

```{r}
fit_brms_simple_hurdle <- 
  brm(
    bf(
      COMM ~ 
        1 + 
        pri_con_fs + 
        grats_spec_fs + 
        age + 
        male + 
        edu, 
      hu ~ 
        1 + 
        pri_con_fs + 
        grats_spec_fs + 
        age + 
        male + 
        edu
    ), 
    data = d, 
    family = hurdle_gamma()
  )
```

```{r}
plot(fit_brms_simple_hurdle, ask = FALSE)
```

```{r}
model_diagnostics(fit_brms_simple_hurdle)
```

```{r}
summary(fit_brms_simple_hurdle)
```

```{r}
conditional_effects(fit_brms_simple_hurdle, ask = FALSE)
```

#### Interaction model

#### Simple hurdle model

```{r}
fit_brms_hurdle_inter <- 
  brm(
    bf(
      COMM ~ 
        1 + 
        pri_con_fs_c * 
        grats_spec_fs_c *
        pri_del_fs_c +
        age + 
        male + 
        edu, 
      hu ~ 
        pri_con_fs_c * 
        grats_spec_fs_c *
        pri_del_fs_c +
        age + 
        male + 
        edu
    ), 
    data = d, 
    family = hurdle_gamma()
  )
```

```{r}
plot(fit_brms_hurdle_inter, ask = FALSE)
```

```{r}
model_diagnostics(fit_brms_hurdle_inter)
```

```{r}
summary(fit_brms_hurdle_inter)
```

```{r}
conditional_effects(fit_brms_hurdle_inter, ask = FALSE)
```

#### Interaction model including deliberation

```{r}
fit_brms <- brm(COMM ~ pri_con_fs_c * grats_spec_fs_c + pri_con_fs_c * pri_del_fs_c + grats_spec_fs_c * pri_del_fs_c + age + male + edu, data = d, family = zero_inflated_poisson())
summary(fit_brms)
```