### 2023-05-12 index.Rmd
minor updates re. formatting

---
title: "Privacy Calculus Experiment"
---

On this website, you can find the companion material for the paper _Likes or Dislikes, Gratifications or Concerns? A One-Week Field Experiment Analyzing the Effects of Popularity Cues on the Privacy Calculus in Online Communication_.

You can either read the paper [in your browser](manuscript.html) or download it as [pdf](manuscript.pdf). You can also find the detailed code and summaries of the [analyses](analyses.html) reported in the paper, as well as [additional analyses](analyses_additional.html) not reported in the paper.

This website is produced directly from the project's [github repository](https://github.com/XMtRa/like_dislike/tree/main). You can download the material and the data, file an issue, or submit a pull request.

![](figures/website/website_translated.png)